Fractional Decomposition Tree Algorithm: A tool for studying the integrality gap of Integer Programs

### 2022-05-05 FDT2020/FDT_revised.tex
main file - fdt revised added
comments almost all done

@@

\begin{abstract}
We present a new algorithm, Fractional Decomposition Tree (FDT) for finding a feasible solution for an integer program (IP) where all variables are binary. FDT runs in polynomial time and is guaranteed to find a feasible integer solution provided the integrality gap is bounded. The algorithm gives a construction for Carr and Vempala's theorem that any feasible solution to the IP's linear-programming relaxation, when scaled by the instance integrality gap, dominates a convex combination of feasible solutions. FDT is also a tool for studying the integrality gap of IP formulations.  We demonstrate that with experiments studying the integrality gap of two problems: optimally augmenting a tree to a 2-edge-connected graph and finding a minimum-cost 2-edge-connected multi-subgraph (2EC). We also give a simplified algorithm, Dom2IP, that more quickly determines if an instance has an unbounded integrality gap. We show that FDT's speed and approximation quality compare well to that of feasibility pump on moderate-sized instances of the vertex cover problem. For a particular set of hard-to-decompose fractional 2EC solutions, FDT always gave a better integer solution than the best previous approximation algorithm (Christofides).
\end{abstract}

>>>>

\usepackage{lineno}
\linenumbers

@@

\begin{abstract}

We present a new algorithm, Fractional Decomposition Tree (FDT) for finding a feasible solution for an integer program (IP) where all variables are binary. FDT runs in polynomial time and is guaranteed to find a feasible integer solution provided the integrality gap is bounded. The algorithm gives a construction for Carr and Vempala's theorem that any feasible solution to the IP's linear-programming relaxation, when scaled by the instance integrality gap, dominates a convex combination of feasible solutions. FDT is also a tool for studying the integrality gap of IP formulations. The upper bound on the integrality gap provided by FDT can be exponentially large. However our experiments demonstrate that FDT can be effective in practice. We study the integrality gap of two problems: optimally augmenting a tree to a 2-edge-connected graph and finding a minimum-cost 2-edge-connected multi-subgraph (2EC). We also give a simplified algorithm, DomToIP, that finds a feasible solution to an IP or concludes that it has unbounded integrality gap. We show that FDT's speed and approximation quality compare well to that of feasibility pump on moderate-sized instances of the vertex cover problem. For a particular set of hard-to-decompose fractional 2EC solutions, FDT always gave a better integer solution than the Best of Many Christofides Algorithm (BOMC).
\end{abstract}

### 2022-05-05 FDT2020/Introduction.tex
comments almost all done

There is an inherent limit to how small $C$ can be for a given IP. The integrality gap for an IP instance is the ratio of the best integer solution to the best solution of the LP relaxation. Any LP-based approximation cannot have an approximation factor $C$ smaller than the integrality gap because there is no feasible solution with an objective value better than a factor of $C$ worse than the optimal solution of the LP relaxation.

@@

Consider an instance of a combinatorial optimization problem $\mathcal{P}$. We can characetrize $\mathcal{P}$ by a set of feasible solutions in $S\in \mathbb{Z}_{\geq 0}^n$ and a cost function $c\in \mathbb{R}_{\geq 0}^n$. In many cases (see \cite{vazirani,shmoyswilliamson}) the set $S$ can decribed as a set of integer points within a polyhedron $P$ (in fact the choice of $P$ need not be unique.) In this case, the integrality gap of $\mathcal{P}$ with respect to $P$ is $\min\{cx:x\in S\}/{\min\{cx:x\in P\}}$. Hence $g(I)$ in Defintion \ref{IGofPoly} is equivalent to the worst-case integrality gap of all instances of a problem with respect to $P(I)$.

@@

@@

One can extend the FDT algorithm for binary IPs into covering\footnote{A covering IP has nonnegative constraint matrix, objective coefficients and right-hand side ($A,b,c$).} $\{0,1,2\}$ IPs by losing a factor $2^{|\spp(x)|}$ on top of the loss for FDT. To eradicate this extra factor, we must treat the coordinate $i$ with $x_i=1$ differently. In Section~\ref{sec:2EC} we focus on the 2-edge-connected multi-subgraph graph problem (2EC) defined above.  Its subtour-elimination LP relaxation is given in (\ref{eq:subtour}).
%}: Given a graph $G=(V,E)$ and $c\in \mathbb{R}^{E}_{\geq 0}$ find a 2-edge-connected multi-subgraph of $G$ with minimum cost. The natural linear programming relaxation for this problem is 
%\begin{equation}
%\min \{cx \; : \; x(\delta(U))\geq 2 \; \text{ for }\emptyset \subset U \subset V, \; x\in [0,2]^E\}
%\end{equation}
For input graph $G$, let $\subtour(G)$ denote the feasible region of this LP. Let $\2ec(G)$ be the convex hull of incidence vectors\footnote{The incidence vector of a 2-edge-connected multi-subgraph $F$, denoted by $\chi^F$ has an element for each edge $e$, with $x_e \in \{0,1,2\}$ indicating the number of times edge $e$ appears in the solution.} of 2-edge-connected multi-subgraphs of graph $G$. Following the definition in (\ref{gapproblem}) have
\begin{equation}
g(\2ec) = \sup_{c\geq 0 , G}\frac{\min_{x\in \2ec(G)} cx}{\min_{x\in \subtour(G)} cx}.
\end{equation}

>>>>

There is an inherent limit to how small $C$ can be for a given IP. The integrality gap for an IP instance is the ratio of the best integer solution to the best solution of the LP relaxation. No LP-based approximation algorithm can have an approximation factor $C$ smaller than the integrality gap because there is not always a feasible solution to the IP with an objective value better than a factor of $C$ times the optimal solution of the LP relaxation.

@@

Consider an instance of a combinatorial optimization problem $\mathcal{P}$. We can characterize $\mathcal{P}$ by a set of feasible solutions in $S\in \mathbb{Z}_{\geq 0}^n$ and a cost function $c\in \mathbb{R}_{\geq 0}^n$. In many cases (see \cite{vazirani,shmoyswilliamson}) the set $S$ can described as a set of integer points within a polyhedron $P$ (in fact the choice of $P$ need not be unique.) In this case, the integrality gap of $\mathcal{P}$ with respect to $P$ is $\min\{cx:x\in S\}/{\min\{cx:x\in P\}}$. Hence $g(I)$ in Definition \ref{IGofPoly} is equivalent to the worst-case integrality gap of all instances of a problem with respect to $P(I)$.

@@

Relatedly, Eisenbrand and Shmonin \cite{eisenbrand} introduced an algorithm that finds the maximum gap between the optimal values of an IP and its LP relaxation over all realizations of vector $b$ (for which the IP is feasible). Their algorithm runs in polynomial-time given $n$ is fixed. In contrast, FDT always runs in time polynomial in $n$, but obtains an upper bound on the integrality gap as opposed to an exact value.

@@

One can extend the FDT algorithm for binary IPs into covering\footnote{A covering IP has nonnegative constraint matrix, objective coefficients and right-hand side ($A,b,c$).} $\{0,1,2\}$ IPs by losing a factor $2$ on top of the loss for FDT. To eradicate this extra factor, we must treat the coordinate $i$ with $x_i=1$ differently. In Section~\ref{sec:2EC} we focus on the 2-edge-connected multi-subgraph graph problem (2EC) defined above.  Its subtour-elimination LP relaxation is given in (\ref{eq:subtour}).
%}: Given a graph $G=(V,E)$ and $c\in \mathbb{R}^{E}_{\geq 0}$ find a 2-edge-connected multi-subgraph of $G$ with minimum cost. The natural linear programming relaxation for this problem is 
%\begin{equation}
%\min \{cx \; : \; x(\delta(U))\geq 2 \; \text{ for }\emptyset \subset U \subset V, \; x\in [0,2]^E\}
%\end{equation}
For input graph $G$, let $\subtour(G)$ denote the feasible region of this LP. Let $\2ec(G)$ be the convex hull of incidence vectors\footnote{The incidence vector of a 2-edge-connected multi-subgraph $F$, denoted by $\chi^F$ has an element for each edge $e$, with $x_e \in \{0,1,2\}$ indicating the number of times edge $e$ appears in the solution.} of 2-edge-connected multi-subgraphs of graph $G$. Following the definition in (\ref{gapproblem}) have
\begin{equation}
g(\2ec) = \sup_{c\geq 0 , G}\frac{\min_{x\in \2ec(G)} cx}{\min_{x\in \subtour(G)} cx}.
\end{equation}

### 2022-05-05 FDT2020/main-section.tex
comments almost all done

\begin{proof}	
	Consider the following linear program which we denote by $\LPC(\ell,x')$. The variables of $\LPC(\ell,x')$ are $\lambda_0$, $\lambda_1$, $x^0$ and $x^1$.
% \cindy{I think you mean that the $\lambda$s are variables, not $\gamma$s.  However the above lemma uses $\gamma$ and then there is a mix of $\lambda$ and $\gamma$ below the LP. Can you please check that you intend to have both types and they are used properly?}
	\begin{align}
	\LPC(\ell,x')\quad\quad& \max\quad \;\lambda_0+\lambda_1\\
	&\;\text{s.t.} \quad Ax^j \geq b\lambda_j & \mbox{ for $j=0,1$} \label{feasibility}\\
	&\;{\color{white}{\text{s.t.}} }\quad 0 \leq x^j \leq \lambda_j &\mbox{ for $j=0,1$}\label{bound}\\
	&\;{\color{white}{\text{s.t.}} }\quad x^0_\ell = 0,\; x^1_\ell =\lambda_1\label{branchcoordinate}\\
	&\;{\color{white}{\text{s.t.}} }\quad x^0 + x^1 \leq x'\label{packing}\\
		&\;{\color{white}{\text{s.t.}} }\quad \lambda_0 + \lambda_1  \leq 1\label{mulitplier-packing}\\
	&\;{\color{white}{\text{s.t.}} }\quad \lambda_0,\lambda_1 \geq 0
	\end{align}

@@

	We now show that $\hat{x}^0$, $\hat{x}^1$, $\gamma_0$, and $\gamma_1$ satisfy all the conditions. Note that conditions (i), (ii), (iii), and (v) are satisfied directly from the  fact that $\bar{x}^0,\bar{x}^1,\gamma_0$ and $\gamma_1$ satisfy these conditions. Thus we only need to show (iv) holds. We need to show that $\gamma_0 \hat{x}^0_j+\gamma_1\hat{x}^1_j\leq x'_j$. If $j=\ell,\ldots,t$, the inequality is satisfied since $\hat{x}^i_j = \bar{x}^i_j$ for $i=0,1$ and property (iv) holds for $\bar{x}^0$ and $\bar{x}^1$. Hence, assume $j\leq \ell-1$. By the property of $x'$ we have $x'_j\in \{0,1\}$. If $x'_j= 0$, then by Constraint (\ref{packing}) we have $\bar{x}^0_j = \bar{x}^1_j=0$. Therefore, $\hat{x}^i_j=0$ for $i=0,1$, so (iv) holds. Otherwise if $x'_j = 1$, then we have
	$\gamma_0\hat{x}^0_j+\gamma_1\hat{x}^1_j\leq \gamma_0+\gamma_1\leq 1\leq x'_j.$ 
	Therefore (v) holds.
\end{proof}

@@

\begin{proof}
	Consider the following LP with variables $\lambda_j$ and $x^j$ for $j=0,1,2$. 
	\begin{align}
	\quad\quad& \max\quad \;\sum_{j=0,1,2}\lambda_j\\
	&\;\text{s.t.} \quad x^j(\delta(U))\geq 2\lambda_j \;& \mbox{ for $\emptyset \subset U \subset V$, and $j=0,1,2$} \label{feasibility2ec}\\
	&\;{\color{white}{\text{s.t.}} }\quad 0 \leq x^j \leq 2\lambda_j\; &\mbox{ for $j=0,1,2$}\label{bound2ec}\\
	&\;{\color{white}{\text{s.t.}} }\quad x^j_e =j\cdot \lambda_j\; &\mbox{ for $j=0,1,2$}\label{branchcoordinate2ec}\\
	&\;{\color{white}{\text{s.t.}} }\quad x^j_f \geq \lambda_j \; &\mbox{ for $f\in E$ where $x_f \geq 1$, and $j=0,1,2$}\label{1edges2ec}\\
	&\;{\color{white}{\text{s.t.}} }\quad x^0 + x^1+x^2 \leq x\label{packing2ec}\\
	&\;{\color{white}{\text{s.t.}} }\quad \lambda_0,\lambda_1,\lambda_2 \geq 0
	\end{align}	Let $x^j$, $\gamma_j$ for $j=0,1,2$ be an optimal solution solution to the LP above. For $j=0,1,2$, if $\gamma_j = 0$, then let $\hat{x}^j=\{j\}^E$, otherwise let $\hat{x}^{j}={x^j}/{\gamma_j}$ for $j=0,1,2$ where $\gamma_j>0$. Observe that  (ii), (iii), (iv), and (v) are satisfied with this choice. We can also show that $\gamma_0+\gamma_1+\gamma_2\geq {1}/{g(\2ec)}$, which means that (i) is also satisfied. The proof is similar to the proof of the claim in Lemma \ref{LPClemma}, but we need to replace each $f\in E$ with $x_f\geq 1$ with a suitably long path to ensure that Constraint (\ref{1edges2ec}) is also satisfied.	
	\begin{claim}\label{CVexists}
		We have $\gamma_0 + \gamma_1+\gamma_2\geq \frac{1}{g(\2ec)}$.
	\end{claim}
	\begin{cproof}
		Suppose for contradiction $\sum_{j=0,1,2}\gamma_j = \frac{1}{g(\2ec)} - \epsilon$ for some $\epsilon >0$. Construct graph $G'$ by removing edge $f$ with $x_f\geq 1$ and replacing it with a path $P_f$ of length $\ceil{\frac{2}{\epsilon}}$. Define $x'_h = x_h$ for each edge $h$ such that $x_h<1$. For each $h\in P_f$ let $x'_h= x_f$ for all $f$ with $x_f\geq 1$. We have $x'\in \subtour(G')$. By Theorem \ref{CV2} there exists $\theta \in [0,1]^k$, with $\sum_{i=1}^{k}\theta_i = 1$ and 2-edge-connected multi-subgraphs $F'_i$ of $G'$ for $i=1,\ldots,k$ such that 
		$\sum_{i=1}^{k}\theta_i \chi^{F'_i}\leq g(\2ec)x'$.

@@

\begin{table}[h!]
	\begin{small}
		\centering
		\begin{tabular}{c c c c}
			\hline
			& number of edges in $T$ & number of links in $L$ & number of instances $(T,L)$\\ \hline
			 & $6$ & $6$ & $100$ \\ 
			 & $14$ & $28$ & $100$ \\ 
			 & $30$ & $120$ & $100$ \\
			 & $62$ & $496$ & $100$ \\ 
			  & $126$ & $2016$ & $100$ \\  \hline \\
		\end{tabular}\caption{Summary of the randomly generated instances of TAP.} \label{tableTAP}
	\end{small}
\end{table}
%\cindy{Please check the edits to the next paragraph carefully to ensure I captured what you were trying to say.}
For all 500 instances in our experiments, running FDT on the LP optimal (fractional extreme point) of the cut LP gave a feasible solution with value at most a factor $\frac{3}{2}$ larger than the LP lower bound.  Such a feasible solution gives an upper bound on the integrality gap $g(\tap)$ of that specific instance of at most $\frac{3}{2}$.
%\cindy{There are 450 instances, so you tested one extreme point (the LP optimal) per instance?  I'm assuming that was the LP optimal.}
%when restricted to the instances considered.
In fact, the integrality gap upper bound derived this way was equal to $\frac{3}{2}$ for only one instance.
%the upper bound we provided by our experiment is at $\frac{3}{2}$.
For 480 instances, the integrality gap upper bound was $\frac{4}{3}$, for 16 instances it was $\frac{6}{5}$, for 2 instances it was $\frac{8}{7}$, for 1 instance it was $\frac{10}{9}$. 
%\arash{it's all correct, I'm running the experiments again to get the running time so these numbers might change.}
\paragraph{Computational comparison between Christofides' algorithm and FDT for 2EC on Carr-Vempala points.}

As described in Section~\ref{sec:2EC-intro}, when the LP optimum for (\ref{eq:subtour}) is a Carr-Vempala point (see Figure~\ref{fig:CVpoint}), the integrality gap of that instance is equal to that of 2EC. So in that sense, these are the hardest fractional points to round to a feasible solution. We now compare the quality of FDT's solution from a Carr-Vempala point to that of Christofides, the best known approximation algorithm\footnote{Recently, \cite{KKO} improved over Christofides algorithm by a small factor.}. Let us remind our reader the Christofides algorithm for TSP: given a graph with edge costs find a minimum spanning tree of the graph.  
%Then, since TSP runs on a complete graph, add a perfect matching on the vertices that have odd degree in the spanning tree to make the (multi)-subgraph 2-edge-connected. If we stop that algorithm without shortcutting, we have a feasible solution for Metric-2EC (given a complete graph, find the minimum-cost 2-edge-connected subgraph, with no multiedges).\arash{The following sentence is not accurate, Metric-TSP is the same as TSP on a non-complete graph where we can take multiple copies of the edges, Metric-2EC (given complete graph with metric costs find the minimum cost 2-edge-connected SUBgraph -no doubling of edges-) is the same as 2EC on a non-complete graph where we can take multiple copies of edges.}\cindy{It sounded like 2EC doesn't assume a complete graph.  But I removed the sentence.  Is the following OK?}\arash{Still not quite accurate. The sentence gives the impression that changing matching with $T$-join is the difference between TSP and 2EC, but its the difference between working with complete graph with metric costs and a non-complete graph where doubling edges is allowed.} 
Next, we add an $O$-join, where $O$ is the set of odd-degree vertices in the spanning tree~\footnote{For graph $G=(V,E)$ and $O\subseteq V$ with $|O|$ even, an $O$-join of $G$ is a subgraph of $G$ that has odd degree on the vertices in $O$ and even degree on vertices in $V\setminus O$. Graphs always have an even number of odd vertices since every edge has two endpoints, giving an even number of endpoints.}. Note that the solutions to the TSP are also 2-edge-connected, which means that Christofides algorithm also provides solutions to the 2EC.
The best-of-many Christofides (BOMC) algorithm~\cite{AKS15} works better than the classic Christofides algorithm in practice~\cite{GW17}.  The BOMC algorithm samples many spanning trees from the LP relaxation (with edges chosen proportionally to the LP value), augments each spanning tree with minimum cost join on the odd degree vertices of the spanning tree, and takes the best solution.

>>>>

\begin{proof}	
	Consider the following linear program which we denote by $\LPC(\ell,x')$. The variables of $\LPC(\ell,x')$ are scalars $\lambda_0$, $\lambda_1$, and vectors $x^0$ and $x^1$.
% \cindy{I think you mean that the $\lambda$s are variables, not $\gamma$s.  However the above lemma uses $\gamma$ and then there is a mix of $\lambda$ and $\gamma$ below the LP. Can you please check that you intend to have both types and they are used properly?}
	\begin{align}
	\LPC(\ell,x')\quad\quad& \max\quad \;\lambda_0+\lambda_1\\
	&\;\text{s.t.} \quad Ax^j \geq b\lambda_j & \mbox{ for $j=0,1$} \label{feasibility}\\
	&\;{\color{white}{\text{s.t.}} }\quad 0 \leq x^j \leq \lambda_j &\mbox{ for $j=0,1$}\label{bound}\\
	&\;{\color{white}{\text{s.t.}} }\quad x^0_\ell = 0,\; x^1_\ell =\lambda_1\label{branchcoordinate}\\
	&\;{\color{white}{\text{s.t.}} }\quad x^0 + x^1 \leq x'\label{packing}\\
		&\;{\color{white}{\text{s.t.}} }\quad \lambda_0 + \lambda_1  \leq 1\label{mulitplier-packing}\\
	&\;{\color{white}{\text{s.t.}} }\quad \lambda_0,\lambda_1 \geq 0
	\end{align}

@@

	We now show that $\hat{x}^0$, $\hat{x}^1$, $\gamma_0$, and $\gamma_1$ satisfy all the conditions. Note that conditions (i), (ii), (iii), and (v) are satisfied directly from the  fact that $\bar{x}^0,\bar{x}^1,\gamma_0$ and $\gamma_1$ satisfy these conditions. Thus we only need to show (iv) holds. We need to show that $\gamma_0 \hat{x}^0_j+\gamma_1\hat{x}^1_j\leq x'_j$. If $j=\ell,\ldots,t$, the inequality is satisfied since $\hat{x}^i_j = \bar{x}^i_j$ for $i=0,1$ and property (iv) holds for $\bar{x}^0$ and $\bar{x}^1$. Hence, assume $j\leq \ell-1$. By the property of $x'$ we have $x'_j\in \{0,1\}$. If $x'_j= 0$, then by Constraint (\ref{packing}) we have $\bar{x}^0_j = \bar{x}^1_j=0$. Therefore, $\hat{x}^i_j=0$ for $i=0,1$, so (iv) holds. Otherwise if $x'_j = 1$, then we have
	$\gamma_0\hat{x}^0_j+\gamma_1\hat{x}^1_j\leq \gamma_0+\gamma_1\leq 1\leq x'_j.$ 
	Therefore (iv) holds.
\end{proof}

@@

\begin{proof}
	Consider the following LP with $\lambda_j$ and $x^j\in \mathbb{R}^E$ for $j=0,1,2$. 
	\begin{align}
	\quad\quad& \max\quad \;\sum_{j=0,1,2}\lambda_j\\
	&\;\text{s.t.} \quad x^j(\delta(U))\geq 2\lambda_j \;& \mbox{ for $\emptyset \subset U \subset V$, and $j=0,1,2$} \label{feasibility2ec}\\
	&\;{\color{white}{\text{s.t.}} }\quad 0 \leq x^j \leq 2\lambda_j\; &\mbox{ for $j=0,1,2$}\label{bound2ec}\\
	&\;{\color{white}{\text{s.t.}} }\quad x^j_e =j\cdot \lambda_j\; &\mbox{ for $j=0,1,2$}\label{branchcoordinate2ec}\\
	&\;{\color{white}{\text{s.t.}} }\quad x^j_f \geq \lambda_j \; &\mbox{ for $f\in E$ where $x_f \geq 1$, and $j=0,1,2$}\label{1edges2ec}\\
	&\;{\color{white}{\text{s.t.}} }\quad x^0 + x^1+x^2 \leq x\label{packing2ec}\\
	&\;{\color{white}{\text{s.t.}} }\quad \lambda_0,\lambda_1,\lambda_2 \geq 0
	\end{align}	Let $x^j$, $\gamma_j$ for $j=0,1,2$ be an optimal solution solution to the LP above. For $j=0,1,2$, if $\gamma_j = 0$, then let $\hat{x}^j=\{j\}^E$, otherwise let $\hat{x}^{j}={x^j}/{\gamma_j}$ for $j=0,1,2$ where $\gamma_j>0$. Observe that  (ii), (iii), (iv), and (v) are satisfied with this choice. We can also show that $\gamma_0+\gamma_1+\gamma_2\geq {1}/{g(\2ec)}$, which means that (i) is also satisfied. The proof is similar to the proof of the claim in Lemma \ref{LPClemma}, but we need to replace each $f\in E$ with $x_f\geq 1$ with a suitably long path to ensure that Constraint (\ref{1edges2ec}) is also satisfied.	
	\begin{claim}\label{CVexists}
		We have $\gamma_0 + \gamma_1+\gamma_2\geq \frac{1}{g(\2ec)}$.
	\end{claim}
	\begin{cproof}
		Suppose for contradiction $\sum_{j=0,1,2}\gamma_j = \frac{1}{g(\2ec)} - \epsilon$ for some $\epsilon >0$. Construct graph $G'$ by removing edge $f$ with $x_f\geq 1$ and replacing it with a path $P_f$ of length $\ceil{\frac{2}{\epsilon}}$. Define $x'_h = x_h$ for each edge $h$ such that $x_h<1$. For each $h\in P_f$ let $x'_h= x_f$ for all $f$ with $x_f\geq 1$. We have $x'\in \subtour(G')$. By Theorem \ref{CV2} there exists $\theta \in [0,1]^k$, with $\sum_{i=1}^{k}\theta_i = 1$ and 2-edge-connected multi-subgraphs $F'_i$ of $G'$ for $i=1,\ldots,k$ such that 
		$\sum_{i=1}^{k}\theta_i \chi^{F'_i}\leq g(\2ec)x'$.

@@

\begin{table}[h!]
	\begin{small}
		\centering
		\begin{tabular}{c c c c}
			\hline
			& number of edges in $T$ & number of links in $L$ & number of instances $(T,L)$\\ \hline
			 & $6$ & $6$ & $100$ \\ 
			 & $14$ & $28$ & $100$ \\ 
			 & $30$ & $120$ & $100$ \\
			 & $62$ & $496$ & $100$ \\ 
			  & $126$ & $2016$ & $100$ \\  \hline \\
		\end{tabular}\caption{Summary of the randomly generated instances of TAP.} \label{tableTAP}
	\end{small}
\end{table}
%\cindy{Please check the edits to the next paragraph carefully to ensure I captured what you were trying to say.}
For all 500 instances in our experiments, running FDT on the LP optimal (fractional extreme point) of the cut LP gave a feasible solution with value at most a factor $\frac{3}{2}$ larger than the LP lower bound.  Such a feasible solution gives an upper bound on the integrality gap $g(\tap)$ of that specific instance of at most $\frac{3}{2}$.
%\cindy{There are 450 instances, so you tested one extreme point (the LP optimal) per instance?  I'm assuming that was the LP optimal.}
%when restricted to the instances considered.
In fact, the integrality gap upper bound derived this way was equal to $\frac{3}{2}$ for only one instance.
%the upper bound we provided by our experiment is at $\frac{3}{2}$.
For 480 instances, the integrality gap upper bound was $\frac{4}{3}$, for 16 instances it was $\frac{6}{5}$, for 2 instances it was $\frac{8}{7}$, for 1 instance it was $\frac{10}{9}$. 
%\arash{it's all correct, I'm running the experiments again to get the running time so these numbers might change.}
\paragraph{Computational comparison between BOMC and FDT for 2EC on Carr-Vempala points.}

As described in Section~\ref{sec:2EC-intro}, when the LP optimum for (\ref{eq:subtour}) is a Carr-Vempala point (see Figure~\ref{fig:CVpoint}), the integrality gap of that instance is equal to that of 2EC. So in that sense, these are the hardest fractional points to round to a feasible solution. We now compare the quality of FDT's solution from a Carr-Vempala point to BOMC. BOMC (and simply Christofides) provide the best known upper bound on $g(\2ec)$\footnote{Recently, \cite{KKO} improved over Christofides algorithm by a small factor.}. Let us remind our reader the Christofides algorithm for TSP: given a graph with edge costs find a minimum spanning tree of the graph.  
%Then, since TSP runs on a complete graph, add a perfect matching on the vertices that have odd degree in the spanning tree to make the (multi)-subgraph 2-edge-connected. If we stop that algorithm without shortcutting, we have a feasible solution for Metric-2EC (given a complete graph, find the minimum-cost 2-edge-connected subgraph, with no multiedges).\arash{The following sentence is not accurate, Metric-TSP is the same as TSP on a non-complete graph where we can take multiple copies of the edges, Metric-2EC (given complete graph with metric costs find the minimum cost 2-edge-connected SUBgraph -no doubling of edges-) is the same as 2EC on a non-complete graph where we can take multiple copies of edges.}\cindy{It sounded like 2EC doesn't assume a complete graph.  But I removed the sentence.  Is the following OK?}\arash{Still not quite accurate. The sentence gives the impression that changing matching with $T$-join is the difference between TSP and 2EC, but its the difference between working with complete graph with metric costs and a non-complete graph where doubling edges is allowed.} 
Next, we add an $O$-join, where $O$ is the set of odd-degree vertices in the spanning tree~\footnote{For graph $G=(V,E)$ and $O\subseteq V$ with $|O|$ even, an $O$-join of $G$ is a subgraph of $G$ that has odd degree on the vertices in $O$ and even degree on vertices in $V\setminus O$. Graphs always have an even number of odd vertices since every edge has two endpoints, giving an even number of endpoints.}. Note that the solutions to the TSP are also 2-edge-connected, which means that Christofides algorithm also provides solutions to the 2EC.
The best-of-many Christofides (BOMC) algorithm~\cite{AKS15} works better than the classic Christofides algorithm in practice~\cite{GW17}.  The BOMC algorithm samples many spanning trees from the LP relaxation (with edges chosen proportionally to the LP value), augments each spanning tree with minimum cost join on the odd degree vertices of the spanning tree, and takes the best solution.
